---
title: "Modestly-weighted logrank test: basic tutorial"
author: "Dominic Magirr"
date: "`r Sys.Date()`"
output: 
  pdf_document:
        fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Basic tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Installation

You can use `devtools::install_github()` to get the package from [GitHub](https://github.com/dominicmagirr/modestWLRT):

```{r eval=FALSE}
install.packages("devtools")
library(devtools)
install_github("dominicmagirr/expectedevents")
install_github("dominicmagirr/modestWLRT")
```


# Load packages

```
library(dplyr)
library(ggplot2)
library(expectedevents)
library(modestWLRT)
```

```{r include=FALSE}
library(dplyr)
library(ggplot2)
devtools::load_all("..")
devtools::load_all("~/AAC/tools/expectedevents")
```

# Perform a mWLRT

## Simulate example data set

You can use the function `delayed_effect_sim` to simulate an example data set from a 2-arm RCT. Survival times on the control arm are exponentially distributed with median `med_c`. Survival times on the experimental arm follow a 2-piece exponential distibution: from time zero up to time `delay` the event rate is `rate_e_1`; thereafter the event rate is `rate_e_2`. Patient recruitment times follow a simple power distribution:

pr(recruited before t) = (t / `rec_period`)^`rec_power`, for t in (0, `rec_period`).

Data cut-off happens at time `max_cal_t`, and any patients still alive have their survival time censored.

```{r}
set.seed(459)
example_data = delayed_effect_sim(n_c = 10,
                                  n_e = 10,
                                  rec_period = 12,
                                  rec_power = 1,
                                  med_c = 15,
                                  rate_e_1 = log(2) / 15,
                                  rate_e_2 = 0.03,
                                  delay = 6,
                                  max_cal_t  = 36)

example_data
```


## Risk table

The function `get_risk_table` takes a data frame produced from `delayed_effect_sim` (or a data frame of the same form) and turns it into a risk table. This tells you how many patients were at risk / had an event / censored on each arm, at each event time.

```{r}
example_risk_table = get_risk_table(example_data)

example_risk_table
```

## Calculate weights

From the risk table, you can calculate the scores / weights from a modestWLRT. The argument `delay` is used to specify how long the scores are kept constant. See the paper

http://arxiv.org/abs/1807.11097

for details.

```{r}
modest_weights = add_weights(example_risk_table, 
                             method = "fixed_c", 
                             delay = 12, 
                             plot_weights = TRUE)
```

```{r}
modest_weights$risk_table
```

```{r fig.cap = "Scores/weights from a modestWLRT."}
modest_weights$p
```

## Test statistics

Given the risk table with the corresponding weights, it is simple to calculate the standardized weighted logrank statistic. Larger values of Z correspond to longer survival times on the experimental arm.

```{r}
get_zs(modest_weights)
```


# Approximate non-centrality parameter / power

To approximate the non-centrality parameter (and therefore power) of a mWLRT, we split the test statistic up into smaller chunks:

$$U = \sum_{k=0}^{\infty} U_k$$

where, for some time interval $t'$,

$$ U_k = \sum_{j:~t_j \in (kt', kt'+t')} ~~w_j\left( d_{0,j} - d_j\frac{n_{0,j}}{n_j}\right).$$

Let $e_k$ and $\theta_k$ denote the number of events and log-hazard ratio, respectively, during the time interval $(kt', kt' + t')$. Approximately, for large $e_k$ and small $\theta_k$,

$$ U_k \sim N(\tilde{w}_k \theta_k I_k, ~\tilde{w}_k^2 I_k), $$
where $I_k = e_k R / (R+1)^2$ for $R:1$ allocation, and $\tilde{w}_k$ denotes an "average" weight during $(kt', kt' + t')$.


When designing the study we make assumptions about survival distributions on each arm, as well as the recruitment distribution. We can use this to calculate 

$$\tilde{w}_k = 1 / S^* \left(\min\left\lbrace kt' + 0.5t', t^* \right\rbrace \right),$$
where $S^*$ denotes the assumed survival distribution for the pooled sample. We also use numerical integration to find

$$E(e_k) = N \times \pi_k$$

where $\pi_k$ denotes the probability of a randomly chosen patient having an event during $(kt', kt' + t')$. 

Putting these pieces together, the standardized logrank statistic $U / \text{var} (U)$ is approximately normally distribution with non-centrality parameter

$$ \sqrt{N} \times \frac{\sum_k \tilde{w}_k \theta_k \pi_k}{\left( \sum_k  \tilde{w}_k^2 \pi_k \right)^{1/2}}.$$

To do this in the `modestWLRT` package, we first specify recruitment assumptions. We assume that

$$ pr(\text{recruited prior to time }r) = (r / R)^k $$
where $R$ denotes the length of the recruitment period.


```{r}
recruitment_1 = list(n_0 = 100, 
                     n_1 = 100, 
                     r_period = 12, 
                     k = 1)

```


Next we specify the survival distributions on the two arms. This is done via piece-wise constant hazards. In this simple example, there are only two pieces. During period 1 (0 -- 6 months), the hazard rate is the same on both arms. During period 2 (6 months onwards), the hazard ratio is 0.5.


```{r}
model_1 = list(change_points = 6, 
               lambdas_0 = c(log(2) / 15, log(2) / 15), 
               lambdas_1 = c(log(2) / 15, log(2) / 30))

```


We want to evaluate the ncp/power for a sequence of pontential values for $t^*$.

```{r}
t_star_seq <- seq(2,36,2)
```


Finally, when calling `ncp_power`, we must also specify the calendar time of data cut-off (`dco`) and the number of pieces to split the log-rank test into (`length_t`). The larger `length_t`, the smaller $t'$.

```{r}
results <- ncp_power(t_star = t_star_seq, 
                     model = model_1,
                     recruitment = recruitment_1,
                     dco = 36,
                     length_t = 50)

```

The output is the approximate non-centrality parameter for each value of $t^*$...

```{r}
plot(t_star_seq, results$ncp, type = 'b')
```

...and corresponding power (by default for one-sided type 1 error rate of 2.5%):

```{r}
plot(t_star_seq, results$power, type = 'b')
```



In this case, the best choice of $t^*$ is about 22 months.









