---
title: "Modestly-weighted logrank test: basic tutorial"
author: "Dominic Magirr"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
        fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Installation

You can use `devtools::install_github()` to get the package from [GitHub](https://github.com/dominicmagirr/modestWLRT):

```{r eval=FALSE}
install.packages("devtools")
library(devtools)
install_github("dominicmagirr/modestWLRT")
```


# Load packages

```
library(dplyr)
library(ggplot2)
library(modestWLRT)
```

```{r include=FALSE}
library(dplyr)
library(ggplot2)
devtools::load_all("..")
```

# Simulate example data set

You can use the function `delayed_effect_sim` to simulate an example data set from a 2-arm RCT. Survival times on the control arm are exponentially distributed with median `med_c`. Survival times on the experimental arm follow a 2-piece exponential distibution: from time zero up to time `delay` the event rate is `rate_e_1`; thereafter the event rate is `rate_e_2`. Patient recruitment times follow a simple power distribution:

$$pr(\text{recruited before t}) = (t / \text{rec_period})^{\text{rec_power}}, ~~t \in (0, \text{rec_period}).$$
Data cut-off happens at time `max_cal_t`, and any patients still alive have their survival time censored.

```{r}
example_data = delayed_effect_sim(n_c = 10,
                                  n_e = 10,
                                  rec_period = 12,
                                  rec_power = 1,
                                  med_c = 15,
                                  rate_e_1 = log(2) / 15,
                                  rate_e_2 = 0.03,
                                  delay = 6,
                                  max_cal_t  = 36)

example_data
```


# Risk table

The function `get_risk_table` takes a data frame produced from `delayed_effect_sim` (or a data frame of the same form) and turns it into a risk table. This tells you how many patients were at risk / had an event / censored on each arm, at each event time.

```{r}
example_risk_table = get_risk_table(example_data)

example_risk_table
```

# Calculate weights

From the risk table, you can calculate the scores / weights from a modestWLRT. The argument `delay` is used to specify how long the scores are kept constant. See the paper

http://arxiv.org/abs/1807.11097

for details.

```{r}
modest_weights = add_weights(example_risk_table, 
                             method = "fixed_c", 
                             delay = 12, 
                             plot_weights = TRUE)
```

```{r}
modest_weights$risk_table
```

```{r fig.cap = "Scores/weights from a modestWLRT."}
modest_weights$p
```

# Test statistics

Given the risk table with the corresponding weights, it is simple to calculate the standardized weighted logrank statistic. Larger values of Z correspond to longer survival times on the experimental arm.

```{r}
get_zs(modest_weights)
```




